Synthesis of Linear Functional Programs

### 2021-05-07 article/article.tex
more

A type system can be formally described through natural deduction -- as a set of rules that determines the concept of deduction for some language~\cite{prawitznd65}.
This comes from the close relation between logic and programming languages, called the Curry-Howard isomorphism or correspondence. A logical proof describes a program, and a program describes a proof.
The main observations made are ``propositions as types'', ``proofs as programs'', ``proof checking as type checking'' \cite{https://www.cs.cmu.edu/~fp/courses/15312-f04/handouts/23-curryhoward.pdf}.
Formalizing our type system, the language i.e. the propositions of the logic are the types themselves, the inference rules are used to construct a proof which corresponds directly to a program, and the correctness of the proof assures the program is well-typed.
The typical natural deduction judgment
\[
    A \textrm{ is a proposition}
\]
is extended to include a \emph{term}
(i.e. to make proofs explicit with a formal definition -- these proofs map directly to a $\lambda$-calculus program as stated by the Curry-Howard correspondence????),
and modified to include a list of hypothesis. 
This extended judgment, called?? a typing judgment, has the form:
\[
    \Gamma \vdash M : A
\]
In the type system context, the list of hypothesis, $\Gamma$, is called \emph{typing environment} (i.e. set of variables and their types),
 the term $M$ is the \emph{program}, and $A$, the natural deduction proposition, is the \emph{type}.
 This judgment can be read -- $M$ has type $A$ in $\Gamma$.

@@

>>>>

\todo{Merge! These 2 paragraphs}
A type system can be formally described through a set of inference
rules that inductively define a judgment of the form $\Gamma \vdash M
: A$, stating that program expression $M$ has type $A$ according to
the \emph{typing assumptions} for variables tracked in $\Gamma$.
\dots

\todo[inline]{Daqui, para tipificar uma função, para Curry-Howard,
  para lógica linear, para proof search, etc\dots Só faz sentido falar de dedução natural no
  contexto de ``lógica'' / Curry-Howard.}

A type system can be formally described through natural deduction -- as a set of rules that determines the concept of deduction for some language~\cite{prawitznd65}.
This comes from the close relation between logic and programming languages, the Curry-Howard isomorphism or correspondence. A logical proof describes a program, and a program describes a proof.
The main observations made are ``propositions as types'', ``proofs as programs'', ``proof checking as type checking'' \cite{https://www.cs.cmu.edu/~fp/courses/15312-f04/handouts/23-curryhoward.pdf}.
Formalizing our type system, the language i.e. the propositions of the logic are the types themselves, the inference rules are used to construct a proof which corresponds directly to a program, and the correctness of the proof assures the program is well-typed.
The typical natural deduction judgment
\[
    A \textrm{ is a proposition}
\]
is extended to include a \emph{term}
(i.e. to make proofs explicit with a formal definition -- these proofs map directly to a $\lambda$-calculus program as stated by the Curry-Howard correspondence????),
and modified to include a list of hypothesis. 
This extended judgment, called?? a typing judgment, has the form:
\[
    \Gamma \vdash M : A
\]
In the type system context, the list of hypothesis, $\Gamma$, is called \emph{typing environment} (i.e. set of variables and their types),
 the term $M$ is the \emph{program}, and $A$, the natural deduction proposition, is the \emph{type}.
 This judgment can be read -- $M$ has type $A$ in $\Gamma$.

@@

% Linear Logic
Linear logic \cite{DBLP:journals/tcs/Girard87} can be thought of as a logic of resources.
Where we were previously able to utilize a variable i.e. a hypothesis as many times as we wanted,
now we'll keep track of the amount available (the ``resources''), and guarantee that every resource is used exactly once, hence the name ``Linear Logic''.
It comes that connectives must be decomposed to get more primitive ones that reflect our resource aware system.
So what would be the $\wedge$ (and) in classical logic, decomposes into its multiplicative
